% Copyright (C) 2014-2016 by Thomas Auzinger <thomas@auzinger.name>

\documentclass[draft,final]{vutinfth} % Remove option 'final' to obtain debug information.

% Load packages to allow in- and output of non-ASCII characters.
\usepackage{lmodern}        % Use an extension of the original Computer Modern font to minimize the use of bitmapped letters.
\usepackage[T1]{fontenc}    % Determines font encoding of the output. Font packages have to be included before this line.
\usepackage[utf8]{inputenc} % Determines encoding of the input. All input files have to use UTF8 encoding.

% Extended LaTeX functionality is enables by including packages with \usepackage{...}.
\usepackage{amsmath}    % Extended typesetting of mathematical expression.
\usepackage{amssymb}    % Provides a multitude of mathematical symbols.
\usepackage{mathtools}  % Further extensions of mathematical typesetting.
\usepackage{microtype}  % Small-scale typographic enhancements.
\usepackage[inline]{enumitem} % User control over the layout of lists (itemize, enumerate, description).
\usepackage{multirow}   % Allows table elements to span several rows.
\usepackage{booktabs}   % Improves the typesettings of tables.
\usepackage{subcaption} % Allows the use of subfigures and enables their referencing.
\usepackage[ruled,linesnumbered,algochapter]{algorithm2e} % Enables the writing of pseudo code.
\usepackage[usenames,dvipsnames,table]{xcolor} % Allows the definition and use of colors. This package has to be included before tikz.
\usepackage{nag}       % Issues warnings when best practices in writing LaTeX documents are violated.
\usepackage{todonotes} % Provides tooltip-like todo notes.
\usepackage{hyperref}  % Enables cross linking in the electronic document version. This package has to be included second to last.
\usepackage[acronym,toc]{glossaries} % Enables the generation of glossaries and lists fo acronyms. This package has to be included last.
%\usepackage{algorithm2e}
% Define convenience functions to use the author name and the thesis title in the PDF document properties.

\usepackage{listings}
\usepackage{upquote}
 

\usepackage{color}
\definecolor{bluekeywords}{rgb}{0.13,0.13,1}
\definecolor{greencomments}{rgb}{0,0.5,0}
\definecolor{redstrings}{rgb}{0.9,0,0}

\lstdefinelanguage{FSharp}%
{morekeywords={let, new, match, with, rec, open, module, namespace, type, of, member, % 
and, for, while, true, false, in, do, begin, end, fun, function, return, yield, try, %
mutable, if, then, else, cloud, async, static, use, abstract, interface, inherit, finally },
  otherkeywords={ let!, return!, do!, yield!, use!, var, from, select, where, order, by },
  keywordstyle=\color{bluekeywords},
  sensitive=true,
  basicstyle=\ttfamily,
	breaklines=true,
  xleftmargin=\parindent,
  aboveskip=\bigskipamount,
	tabsize=4,
  morecomment=[l][\color{greencomments}]{///},
  morecomment=[l][\color{greencomments}]{//},
  morecomment=[s][\color{greencomments}]{{(*}{*)}},
  morestring=[b]",
  showstringspaces=false,
  literate={`}{\`}1,
  stringstyle=\color{redstrings},
}







\newcommand{\authorname}{Bernhard Rainer} % The author name without titles.
\newcommand{\thesistitle}{Interactive Shape Detection in Out-of-Core Point-Clouds for assisted User Interactions} % The title of the thesis. The English version should be used, if it exists.

% Set PDF document properties
\hypersetup{
    pdfpagelayout   = TwoPageRight,           % How the document is shown in PDF viewers (optional).
    linkbordercolor = {Melon},                % The color of the borders of boxes around crosslinks (optional).
    pdfauthor       = {\authorname},          % The author's name in the document properties (optional).
    pdftitle        = {\thesistitle},         % The document's title in the document properties (optional).
    pdfsubject      = {Subject},              % The document's subject in the document properties (optional).
    pdfkeywords     = {a, list, of, keywords} % The document's keywords in the document properties (optional).
}

\setpnumwidth{2.5em}        % Avoid overfull hboxes in the table of contents (see memoir manual).
\setsecnumdepth{subsection} % Enumerate subsections.

\nonzeroparskip             % Create space between paragraphs (optional).
\setlength{\parindent}{0pt} % Remove paragraph identation (optional).

\makeindex      % Use an optional index.
\makeglossaries % Use an optional glossary.
%\glstocfalse   % Remove the glossaries from the table of contents.

% Set persons with 4 arguments:
%  {title before name}{name}{title after name}{gender}
%  where both titles are optional (i.e. can be given as empty brackets {}).
\setauthor{}{\authorname}{BSc.}{male}
\setadvisor{Pretitle}{Michael Wimmer}{Posttitle}{male}

% For bachelor and master theses:
%\setfirstassistant{Pretitle}{Forename Surname}{Posttitle}{male}
%\setsecondassistant{Pretitle}{Forename Surname}{Posttitle}{male}
%\setthirdassistant{Pretitle}{Forename Surname}{Posttitle}{male}

% For dissertations:
\setfirstreviewer{Pretitle}{Forename Surname}{Posttitle}{male}
\setsecondreviewer{Pretitle}{Forename Surname}{Posttitle}{male}

% For dissertations at the PhD School and optionally for dissertations:
\setsecondadvisor{Pretitle}{Forename Surname}{Posttitle}{male} % Comment to remove.

% Required data.
\setaddress{Heigerleinstraße 53 8, 1170 Wien}
\setregnumber{0828592}
\setdate{1}{03}{2017} % Set date with 3 arguments: {day}{month}{year}.
\settitle{\thesistitle}{\thesistitle} % Sets English and German version of the title (both can be English or German).
%\setsubtitle{Optional Subtitle of the Thesis}{Optionaler Untertitel der Arbeit} % Sets English and German version of the subtitle (both can be English or German).

% Select the thesis type: bachelor / master / doctor / phd-school.
% Bachelor:
%\setthesis{bachelor}
%
% Master:
\setthesis{master}
\setmasterdegree{dipl.} % dipl. / rer.nat. / rer.soc.oec. / master
%
% Doctor:
%\setthesis{doctor}
%\setdoctordegree{rer.soc.oec.}% rer.nat. / techn. / rer.soc.oec.
%
% Doctor at the PhD School
%\setthesis{phd-school} % Deactivate non-English title pages (see below)

% For bachelor and master:
\setcurriculum{Visual Computing}{Visual Computing} % Sets the English and German name of the curriculum.

% For dissertations at the PhD School:
\setfirstreviewerdata{Affiliation, Country}
\setsecondreviewerdata{Affiliation, Country}


\begin{document}

\frontmatter % Switches to roman numbering.
% The structure of the thesis has to conform to
%  http://www.informatik.tuwien.ac.at/dekanat

\addtitlepage{naustrian} % German title page (not for dissertations at the PhD School).
\addtitlepage{english} % English title page.
\addstatementpage

%\begin{danksagung*}
%\todo{Ihr Text hier.}
%\end{danksagung*}

\begin{acknowledgements*}

This thesis would not have been possible without the help of a handful engaged people. 
My first and foremost gratitude goes out to the VRVis Zentrum für Virtual Reality und Visualisierung for providing me with the opportunity to serve an internship and in consequence conduct this thesis. Within the VRVis, I would like to thank the Semantic Modeling and Acquisition (SMAQ) group for integrating me into their team, both on a professional level and amicably. A special thank you goes out to my Aardvark, and F\# Gurus Harald Steinlechner, Georg Haaser, and Attila Szabo, whose help and expertise eased the execution of this thesis enormously. 
I would also like to thank Stefan Maierhofer, leader of the SMAQ group, and my project manager Michael Schwärzler for their ideas and continuous support. 
\\
A big thank you goes out to Michael Wimmer from the Technische Universität Wien for his supervision on not only this thesis but all preceding courses and practica. 

Thanks to Lisa Kellner, who accompanied me through this thesis on a daily basis as she sat next to me and continuously provided me with feedback on my work. 

Thanks to Phillip Erler, a fellow diploma student, for all the exchange of knowledge on both our thesis’s. 

Last but not least, I would like to thank my friends and family that endured my ongoing talks about my thesis for the bigger part of a year and encouraged me to pursue an academic career.  

\end{acknowledgements*}

%\begin{kurzfassung}
%\todo{Ihr Text hier.}
%\end{kurzfassung}

\begin{abstract}

This thesis presents an interactive method of shape detection for out-of-core point clouds. By utilizing the results of the shape detection, improvements are proposed to existing two-dimensional interaction metaphors. Usually, shape detection is performed on the whole point cloud. Thus results are obtained after an extended period of time. Instead, our approach splits the point cloud into chunks of neighboring points that can be processed efficiently in near real-time. The user can select such a region of interest for it to be segmented and is provided with feedback within a fraction of a second, thus making the detected geometry usable almost immediately. The resulting geometries, called primitive shapes, are used to improve existing two-dimensional interaction metaphors. Lasso Selection, as well as Volumetric Brush Selection, benefit from the use of primitive shapes to support the interaction, such that only points are selected that lie roughly follow the curvature of the shape. Thus, the selection of unwanted points in the foreground or background is reduced. Point Picking is improved by the use of a primitive shape as support, as only points are pickable that also belong to the detected primitive shape.\textit{Local Level-of-Detail Increment} is a novel interaction that increases the level-of-detail along a primitive shape, such that additional points that would otherwise be culled are rendered as well. 
\\

As modern point clouds can contain billions of points and the memory capacity of consumer PCs is usually insufficient to store all points at all time, a level-of-detail data structure is used to store the point cloud on the hard disc and data is loaded into memory only on use. This thesis uses a functional octree as data structure to store all points and create a level-of-detail representation of the point cloud. The level-of-detail representation of the point cloud, based on the camera's relative position, is used for interactions and rendering of the point cloud, such that distant regions are presented to the user at lower point resolutions. 
This work was enabled by the Competence Centre VRVis. 

\end{abstract}

% Select the language of the thesis, e.g., english or naustrian.
\selectlanguage{english}

% Add a table of contents (toc).
\tableofcontents % Starred version, i.e., \tableofcontents*, removes the self-entry.

% Switch to arabic numbering and start the enumeration of chapters in the table of content.
\mainmatter


\input{Introduction/Introduction.tex} 
\input{Related_Work/Related_work.tex}
\input{Octree/Octree.tex}
\input{Shape_Detection/Shape_Detection.tex}
\input{Shape_Detection/Shape_Clustering.tex}
\input{System_Design/System_Design.tex}
\input{Implementation/Implementation.tex}
\input{Results/Results.tex}
\input{Conclusion/Conclusion.tex}

\backmatter

% Use an optional list of figures.
\listoffigures % Starred version, i.e., \listoffigures*, removes the toc entry.

% Use an optional list of tables.
\cleardoublepage % Start list of tables on the next empty right hand page.
\listoftables % Starred version, i.e., \listoftables*, removes the toc entry.

% Use an optional list of alogrithms.
\listofalgorithms
\addcontentsline{toc}{chapter}{List of Algorithms}

% Add an index.
\printindex

% Add a glossary.
\printglossaries

% Add a bibliography.
\bibliographystyle{alpha}
\bibliography{intro}

\end{document}